---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity.data<-read.csv('~/Documents/R Programming/Reproducible Research/activity.csv')
activity.data$date<-as.Date(as.character(activity.data$date),'%Y-%m-%d')
```
## What is mean total number of steps taken per day?  

**1. Calculate the total number of steps taken per day**  
  
```{r}
steps.perday<-aggregate(steps~date,activity.data,sum,na.rm=F)
print(steps.perday)
```
  
**2. Make a histogram of the total number of steps taken each day.**  

```{r}
hist(steps.perday$steps,breaks=10,main='Total number of steps taken daily',xlab='Number of steps',ylab='Number of Days')
```  
  
**3. Calculate and report the mean and median of the total number of steps taken per day**  

```{r}
steps.perday.mean<-mean(steps.perday$steps)
steps.perday.median<-median(steps.perday$steps)
```  
  
**MEAN = **`r format(steps.perday.mean,scientific=F)` (Average number of daily steps)  
**MEDIAN= **`r steps.perday.median` (Median of the number of daily steps)  


## What is the average daily activity pattern?  

**1. Make a time serie plot...**  

```{r}
steps.interval<-aggregate(steps~interval,activity.data,na.rm=F,mean)
with(steps.interval,plot(interval,steps,type='l',xlab='Interval',ylab='Number of steps',main='Average number of steps (all days) per interval'))
max.interval<-steps.interval[which.max(steps.interval$steps),1]
```
  
**2. Which 5-minute interval, on average across all days, contains the maximum number of steps**  

```{r}

max.interval<-steps.interval[which.max(steps.interval$steps),1]
```
  
On average, interval **`r max.interval`** has the most steps with **`r round(max(steps.interval$steps))`** steps  
  

## Imputing missing values  
  
**1. Number of missing values**  
```{r}

number.NA<-sum(is.na(activity.data))
mean.NA<-mean(is.na(activity.data))
```
  
There is **`r number.NA`** missing values which represents **`r round(100*mean.NA)` %** of missing values  

**2. Devise a strategy for filling the missing values**  

As suggested in the instruction, I decided to associate all NA values to the mean of that interval.  
**3. Create a new dataset with missing data filed**  

```{r}
imputed.activity.data<-merge(activity.data,steps.interval, by='interval')
NA.steps<-which(is.na(imputed.activity.data$steps.x))
imputed.activity.data$steps.x[NA.steps]<-round(imputed.activity.data$steps.y[NA.steps])
imputed.activity.data<-imputed.activity.data[,1:3]
imputed.daily.steps<-aggregate(steps.x~date,imputed.activity.data,sum)
```
  
**4. Make histogram**  

```{r}

mean.daily.steps<-mean(imputed.daily.steps$steps.x)
median.daily.steps<-median(imputed.daily.steps$steps.x)
hist(imputed.daily.steps$steps.x,col='red',breaks=10,main='Total number of steps taken daily (NA replaced by interval mean)',xlab='Number of steps',ylab='Number of Days')

```
  
After replacing the the missing values, we realise that the difference is not signigicant. We could've ignored the NA datas.  

MEAN Diff% =(Ignored-Imputed)/Ignored = `r  round(steps.perday.mean-mean.daily.steps)/steps.perday.mean*100 ` %  

MEDIAN Diff% =(Ignored-Imputed)/Ignored = `r round(steps.perday.median-median.daily.steps)/steps.perday.median*100` %  
  
  
## Are there differences in activity patterns between weekdays and weekends?
  
**1. Create new Factor variable**  

```{r}
imputed.activity.data$weekday<-as.factor(weekdays(imputed.activity.data$date))
weekday.name<-c('Monday','Tuesday','Wednesday','Thursday','Friday')
imputed.activity.data$weekend<-ifelse(imputed.activity.data$weekday %in% weekday.name,'Weekday','Weekend')
imputed.interval.mean<-aggregate(steps.x~interval+weekend,imputed.activity.data,mean)
```
  
**2. Panel plot**
  
```{r}
library(lattice)
with(imputed.interval.mean,xyplot(steps.x~interval|weekend,type='l',ylab='Avg number of steps',xlab='Interval',layout=c(1,2)))
```
  
The pattern is a little different at the steps per interval during the weekend seem more distributed
